Contextual Identity: Towards Usable Privacy

### 2013-02-28 paper.tex
reorg first page

\title{Contextual Identity: A Framework for Usable Privacy}

\author{Monica Chew,Sid Stamm\\
\tt{\{mmc,sid\}@mozilla.com}}
%\institute{Mozilla}
\maketitle
\section{Introduction}
\begin{quote}I am large, I contain multitudes. --- Walt Whitman,
\begin{em}Song of Myself\end{em} \end{quote}

People reveal different aspects of themselves depending on context. At any
given time, they may be acting as a friend, relative, spouse, co-worker,
acquaintance, or stranger. This observation is not new; Erving Goffman called
this impression management, and Carl Jung called it persona
theory~\cite{goffman,jung}. In this short paper, we call it contextual identity.
%identity as social construct?

Helen Nissenbaum has long argued that privacy violations come not from the
simple sharing of personal information, but rather from sharing that
information in a way that violates social norms, or in the wrong
context~\cite{nissenbaum}. The desire for spontaneous, positive human
interaction often leads to sharing personal information. For example, take
PatientsLikeMe~\cite{patientslikeme}. This site functions as a support group
and as a place for people to share health information, such as drug regimens,
treatment history, side-effects, and more. Though this information is
traditionally considered extremely sensitive, the users of this site clearly
value the benefits of sharing: longitudinal graphs of their own health history,
building community with people who share their disease. Further, sharing
personal information does not mean that the users don't value
privacy~\cite{boyd2}. Selling user data originally intended for one purpose for
another (e.g., to insurance companies to discover uninsurable patients) would
be a gross violation of users' privacy expectations.

%this story is a positive story to shock security folk with the breadth and
%depth of personal information that ordinary users will share
% instead we could lead with a shocking, negative story like how gay youth get
% killed when they're outed

In offline environments, managing contextual identities is more intuitive than
in online environments. There are cues to help you determine where you are,
who your intended audience is, how many people will overhear you, and how
likely your information is to be re-broadcast in a different context. It is
also true that in offline environments, humans don't have perfect memories and
will eventually forget.  However, with the advent of increasingly vast
searching, indexing, and archiving capabilities, one cannot rely on forgetting
in online environments~\cite{delete}. Thus, managing contextual identities
online has become increasingly difficult and fraught with mistakes.

One possible solution to this problem is never to post personal information
online, or at least to consider the entire world might be the audience.
However, this solution defies human nature. Given that there are 1 billion
active Facebook users and that 30-40\% of spoken communication is devoted to
informing others about ourselves, humans are bound to fall short of these
guidelines~\cite{tamir,fbusers}. The security and privacy community should
recognize this and develop techniques to meet users' needs rather than
expecting human nature to change.

@@

Barth et al.~provide a formal model for contextual integrity of user
data~\cite{barth}. The contextual integrity model focuses on the flow of
personal information in terms of knowledge states held by agents rather than on
user identity, and works well for modeling privacy legislation and policy. Two
drawbacks of this work are that it is incapable of modeling information that is
about more than one person, and it is the type of information rather than the
content of the information that determines whether or not integrity is
violated. In the context of social networks, information is rarely ever about a
single person, and the tone of the message is as important to user
happiness as the content of the message.

@@

\begin{comment}TODO: figure out how to cite this. Why is this short
paper's model worth pursuing? It's simpler, and focuses on the user rather
than data.
\end{comment}

@@

\section{Contextual identity violations}
\label{sec:examples}
\subsection{Redistributing information out of context}
Re-distributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals and more to the
user's activity stream.  Because of poor opt-out and lack of visibility into
what was being published, user outcry was immediate~\cite{mccarthy,nakashima}.
In December 2007, Google Reader exposed RSS feeds of user-marked news stories
to the user's Google Talk contacts\footnote{Google Talk contacts are everyone
with whom a user has chatted, which might include co-workers, supervisors, and
friends}. Although this feed was always public, prior to this launch it was not
discoverable, leading to a bad experience for many users~\cite{helft}. In
February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts which users used to treat as private
messages, before the advent of ``Like'' and comment buttons, to the attention
of an audience that the user never anticipated~\cite{timeline}.

\subsection{Unsatisfiable policies}
\label{sec:policies}
Some service providers have policies which preclude isolating multiple
contextual identities. For example, Facebook and Google have a ``Real Names'' policy, which
requires users to register for accounts with their legal
name~\cite{fb_names,google_names}.  These policies presume the users can be one
person to everyone. These policies ignore that community-building happens in
many different contexts, that users have many legitimate reasons for presenting
different identities in different contexts, and that users don't necessarily
want those identities to be linked. For example, disallowing avatar handles as
a primary identifier makes building a gaming community difficult. 
Linking a gaming identity to a professional identity has already caused
problems for 2012 Maine Senate candidate and World of Warcraft gamer Colleen
Lachowicz~\cite{maine}.

@@

@@

\subsection{Federated login}
\label{sec:login}
Facebook Connect is a login platform that allows third-party websites to
authenticate users using their Facebook identity~\cite{fb_connect}. Because it
is against the terms of service to have multiple Facebook accounts, using
Facebook Connect has the unwanted side-effect of linking multiple contextual
identities.

Similar to section~\ref{sec:policies}, many federated login systems simply
don't have the ability to allow users multiple identifiers.
Other federated login systems have support for multiple identifiers, but
this feature can be poorly implemented or difficult to discover, leading to
low use.
Some login platforms and protocols, such as OAuth and BrowserID,
do not suffer from this policy or design error~\cite{browserid,oauth},
giving a user better control over which contextual identity to assert.

@@

\subsection{Inferring contextual identities}
Accurately detecting the user's current contextual identity, whether
heuristically or through asking the user, is paramount to helping the user
manage it.
Although a user may subconsciously be aware of their contextual identities,
expecting users to enumerate their identities and manually curate them is
probably too burdensome. Firefox has a Profile Manager, but it is so difficult to use that it will be removed.\footnote{See \url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}}.

@@

\subsubsection{Remembering to Forget}
Humans don't have perfect memories, and perhaps social
networks should not either~\cite{delete}. Many users already engage in manual auditing and
deletion of old posts~\cite{fbtips2}. A better solution would be to build tools
that let the user manage this more easily through an API: both Twitter and Facebook provide
APIs for deletion\footnote{At the time of this writing, Google Plus offers a
read-only API}. One way to reduce the likelihood of linked contextual
identities would be to provide applications that expire posts after some time.
Deletion and expiration also reduce the likely audience especially in the case of public posts

\begin{comment}
\cite{viegas}blogger's expectation of privacy
\end{comment}

@@

\section{Discussion}
People have multiple contextual identities. Sharing personal information in
those contexts is a great way to foster positive human interaction.  We hope
that contextual identity will serve as a framework for both developers to
understand their users' privacy needs, and to develop tools that allow users to
act in their contextual identities in a spontaneous, relaxed, and interesting
way.

>>>>

\title{Contextual Identity: Towards Usable Privacy}

\author{Monica Chew,Sid Stamm\\
\tt{\{mmc,sid\}@mozilla.com}}
%\institute{Mozilla}
\maketitle

\begin{abstract}
We examine \textit{contextual identity}, a notion that individuals reveal
different aspects of themselves depending on context. At any given time, they
may act as a friend, relative, spouse, co-worker, aquaintance, or stranger. We
analyze contextual identity from the perspective of user choice and control,
survey contextual identity violations, suggest possible mitigations, and
propose new resesarch directions to empower users to have better control over their contextual identities.
\end{abstract}
\section{Introduction}
\begin{quote}I am large, I contain multitudes. --- Walt Whitman,
\begin{em}Song of Myself\end{em} \end{quote}

The desire for spontaneous, positive human interaction often requires sharing
personal information. Sharing information can obviously lead to loss of
privacy, but the act of sharing alone does not negate expectation of privacy~\cite{boyd2}.
Helen Nissenbaum has long argued that privacy violations come not from the
simple sharing of personal information, but rather from sharing that
information in a way that violates social norms, or in the wrong
context~\cite{nissenbaum}.

\begin{verbatim}
NEED NEW EXAMPLE
gay teen joins facebook group
megan meier
amanda todd
tyler clementi
http://www.sync-blog.com/sync/2012/10/gay-students-outed-by-facebook-privacy-loophole.html
http://books.google.ca/books?id=RRUkLhyGZVgC&pg=PA199&lpg=PA199&dq=%22The+days+of+you+having+a+different+image+for+your+work+friends+or+co-workers+and+for+the+other+people+you+know+are+probably+coming+to+an+end+pretty+quickly,%22&source=bl&ots=LRUZf_pd9m&sig=oHHi0qhlHh9prk94gCQFXItjGIg&hl=en&sa=X&ei=ma19UIOhAdOEqQHE2oCADA&ved=0CD8Q6AEwBA#v=onepage&q=%22The%20days%20of%20you%20having%20a%20different%20image%20for%20your%20work%20friends%20or%20co-workers%20and%20for%20the%20other%20people%20you%20know%20are%20probably%20coming%20to%20an%20end%20pretty%20quickly%2C%22&f=false
\end{verbatim}

In offline environments, managing contextual identities is more intuitive than
in online environments. There are cues to help you determine where you are,
who your intended audience is, how many people will overhear you, and how
likely your information is to be re-broadcast in a different context. It is
also true that in offline environments, humans don't have perfect memories and
will eventually forget.  However, with the advent of increasingly vast
searching, indexing, and archiving capabilities, one cannot rely on forgetting
in online environments~\cite{delete}. Managing contextual identities
online has become increasingly difficult and fraught with mistakes.

@@

The notion that individuals inhabit multiple personas has existed in multiple
fields for many years; Erving Goffman called this impression management, and
Carl Jung called it persona theory~\cite{goffman,jung}.

Barth, Datta, Mitchell, and Nissenbaum provide a formal model for contextual
integrity of user data~\cite{barth}. The contextual integrity model focuses on
the flow of personal information in terms of knowledge states held by agents
acting in given roles, and works well for modeling privacy legislation such as
HIPPA and COPPA. Two drawbacks of this work are that it is incapable of
modeling information where the subject involves more than one person, and it
only takes the type of information into account rather than the content or tone
of the information.  Thus, the event ``Bob told Alice that Charlie cooked a
delicious dinner'' is out of scope of the model.  In the context of social
networks, information is rarely ever about a single person, and the tone of the
message is as important to user happiness as the content of the message.

@@

@@

\section{Contextual identity violations}
\label{sec:examples}
\subsection{Redistributing information out of context}
Re-distributing information out of its original context often leads to
embarrassment~\cite{nissenbaum}.  In November 2007, Facebook Beacon allowed
third-party sites to publish purchases, travel bookings, movie rentals and more to the
user's activity stream.  Because of poor opt-out and lack of visibility into
what was being published, user outcry was immediate~\cite{mccarthy,nakashima}.
In December 2007, Google Reader exposed RSS feeds of user-marked news stories
to the user's Google Talk contacts, which includes everyone
with whom a user has chatted whether they be co-workers, supervisors, or
friends. Although this feed was always public, prior to this launch it was not
discoverable, leading to a bad experience for many users~\cite{helft}. In
February 2010, Google Buzz launched, a product that
exposed the user's most frequent Gmail and Google Talk contacts and their
publicly available (through previously less discoverable) news and photos.  The
combination of exposing contacts (which could include abusive ex-husbands,
co-workers, and friends), as well as aggressively linking photos and news
items, resulted in a huge backlash~\cite{fugitivus,buzz}. In September 2012,
Facebook imported old wall posts from 2008 into the new Timeline interface.
Although wall posts were always visible from profile pages, the new Timeline
interface brought old wall posts which users used to treat as private
messages, before the advent of ``Like'' and comment buttons, to the attention
of an audience that the user never anticipated~\cite{timeline}.

\subsection{Unsatisfiable policies}
\label{sec:policies}
Some service providers have policies which preclude isolating multiple
contextual identities. For example, Facebook and Google have a ``Real Names''
policy, which requires users to register for accounts with their legal
name~\cite{fb_names,google_names}.  These policies ignore that
community-building happens in many different contexts, that users have many
legitimate reasons for presenting different identities in different contexts,
and that users don't necessarily want those identities to be linked. For
example, disallowing avatar handles as a primary identifier makes building a
gaming community difficult.  Linking a gaming identity to a professional
identity has already caused problems for 2012 Maine Senate candidate and World
of Warcraft gamer Colleen Lachowicz~\cite{maine}.

@@

Even worse, many of these providers now serve as login platforms for external
sites, spreading these unsatisfiable policies outside the service providers
themselves.  Facebook Connect is a login platform that allows third-party
websites to authenticate users using their Facebook identity~\cite{fb_connect}.
Because it is against the terms of service to have multiple Facebook accounts,
using Facebook Connect has the unwanted side-effect of linking multiple
contextual identities.

Other federated login systems have support for multiple identifiers, but this
feature can be poorly implemented or difficult to discover, leading to low use.
Some login platforms and protocols, such as OAuth and BrowserID, do not suffer
from this policy or design error~\cite{browserid,oauth}, giving a user better
control over which contextual identity to assert.

@@

@@

\subsection{Inferring contextual identities}
Accurately detecting the user's current contextual identity, whether
heuristically or through asking the user, is paramount to helping the user
manage it.
Although a user may subconsciously be aware of their contextual identities,
expecting users to enumerate their identities and manually curate them is
probably too burdensome. Firefox has a Profile Manager, but it is so difficult to use that it will be removed.\footnote{See \url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}.}

@@

\subsubsection{Remembering to Forget}
Humans don't have perfect memories, and perhaps social networks should not
either~\cite{delete}. Many users already engage in manual auditing and deletion
of old posts~\cite{fbtips2}. A better solution would be to build tools that let
the user manage this more easily through an API: both Twitter and Facebook
provide APIs for deletion\footnote{At the time of this writing, Google Plus
offers a read-only API.}. One way to reduce the likelihood of linked contextual
identities would be to provide applications that expire posts after some time.

@@

\section{Discussion}
People have multiple contextual identities. Sharing personal information in
appropriate contexts fosters positive human interaction.  We hope that
contextual identity will serve as a useful notion for developers to understand
their users' privacy needs, and to develop tools that allow users to act in
their contextual identities in a spontaneous, relaxed, and interesting way.
